Boundary Labeling for annotated documents

### 2017-12-27 LaTeX/Paper.tex
Paper: Added Section 1.2 - Related Work, and changed parts of Chapter 1 to fit the new section. Program: No changes.

\setaddress{TODO: ADDRESS!}%TODO
\setregnumber{1125755}
\setdate{25}{09}{2017}

@@

\chapter{Introduction}
Whenever additional information neets to be inserted into an existing document without altering the original text, we can make use of annotations. They usually take the form of footnotes, which require only a minimal reference in the main text, and are used for a variety of reasons - for example, to provide additional information that would hinder the text's flow if inserted directly, or as a result of a commenting tool that is used for communication between an author and their editor.
If a more obvious connection between the text and the referenced content is required, for example when lengthy comments are added, or if a change-tracking tool is used, the reference is  placed to the side of the document and visibly connected to the part of the text it is referring to. This style of attotation is easily implemented on virtual documents, however if the annotations need to be included in a printed version, there are several issues that arise regarding readability of the final product and ambiguity of text-annotation assignments.

In this paper, we will look at ways to use Boundary Labeling for this problem, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing. (See also \cite{Bekos2007})

@@

While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text. Generally, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results. %Beispiele von suboptimalen Lösungen einfügen? (Bilder)
However, the papers that do discuss boundary labeling in text offer interesting contributions.

In the paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different styles of drawing these connecting paths, and came to the conclusion that paths without bends are easier to follow. %Bild einfügen?
However, most solutions proposed in that paper do not consider whether a path overlaps with text or not, which results in a decrease in readability.

The thesis by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer paths, and forces curves, but doesn't obscure any part of the text.
%Siehe auhc andere Papers - Motivation! (Warum will ich mir dieses Thema ansehen)

\section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, Euclidean space.
To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)

@@

\emph{Polylines} are a connected series of \emph{line segments}. Line segments are straight lines that contain each point between their starting and end point. 
\emph{Labels} hold additional information and are represented as boxes containing this information. They are usually placed in the \emph{Label area} which is a rectangular area designated to hold labels. It is located next to of the bigger \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to. The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area. \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination. \emph{S-Segments} are not required to have any particular orientation, and simply connect their start and ending points in a straight line.
The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref{fig:term} would be classified as an OPOPO-Leader. \change{(Farben angepasst, hoffentlich ist es jetzt in Schwarz-Weiß besser zu erkennen)}
The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity

@@

>>>>

\setaddress{Scherzergasse 10/8, 1020 Wien}
\setregnumber{1125755}
\setdate{25}{09}{2017}%TODO: Update

@@

\chapter{Introduction}
Whenever additional information neets to be inserted into an existing document without altering the original text, we can make use of annotations. They usually take the form of footnotes, which require only a minimal reference in the main text, and are used for a variety of reasons - for example, to provide additional information that would hinder the text's flow if inserted directly, or as a result of a commenting tool that is used for communication between an author and their editor.
If a more obvious connection between the text and the referenced content is required, for example when lengthy comments are added, or if a change-tracking tool is used, the reference is often placed to the side of the document and visibly connected to the part of the text it is referring to. This style of annotation is easily implemented on virtual documents, since they can be hidden on demand, however if the annotations need to be included in a printed version, there are several issues that arise regarding readability of the final product and ambiguity of text-annotation assignments.

In this thesis, we will look at ways to use Boundary Labeling for this problem, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing. (See also \cite{Bekos2007})

@@

\section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
Boundary Labeling (or equivalent concepts) can be applied to a space with different geometry or more dimensions, but this thesis will only concern itself with two-dimensonal, Euclidean space.
To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)

@@

\emph{Polylines} are a connected series of \emph{line segments}. Line segments are straight lines that contain each point between their starting and end point. 
\emph{Labels} hold additional information and are represented as boxes containing this information. They are usually placed in the \emph{Label area} which is a rectangular area designated to hold labels. It is located next to of the bigger \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to. The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area. \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination. \emph{S-Segments} are not required to have any particular orientation, and simply connect their start and ending points in a straight line.
The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref{fig:term} would be classified as an OPOPO-Leader.
The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity

@@

\section{Related Work}

While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text. Generally, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results. %Beispiele von suboptimalen Lösungen einfügen? (Bilder)

The concept of boundary labeling is introduced in \cite{Bekos2007}, where several approaches with varying leader types are discussed, along with any problems that arise. 
However, the papers that do discuss boundary labeling in text offer interesting contributions.

The paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different styles of drawing these connecting paths, and came to the conclusion that paths without bends are easier to follow. %Bsp-Bild einfügen!
However, most solutions proposed in that paper do not consider whether a path overlaps with text or not, which results in a decrease in readability.

The thesis by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer paths, and forces curves, but doesn't obscure any part of the text.
